Stats

### 2023-03-10 stats/analysis.Rmd
prepare repo for revision

    # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
    # the smaller the value, the more central the position
    symmetricPosition = case_when(
                                  targetPosition == "5" | targetPosition == "6" | targetPosition == "box3" ~ 1, 
                                  targetPosition == "4" | targetPosition == "7" | targetPosition == "box2" | targetPosition == "box4" ~ 2,
                                  targetPosition == "3" | targetPosition == "8" | targetPosition == "box1" | targetPosition == "box5" ~ 3, 
                                  targetPosition == "2" | targetPosition == "9" ~ 4,
                                  targetPosition == "1" | targetPosition == "10" ~ 5, 
                       ),

@@

```{r vali_data}
vali <- testtrials %>% 
  filter(studytype == "vali" & (ageInYears < 6 | ageInYears == 18)) %>% 
  mutate(ageInYears = factor(ageInYears, levels = c("3", "4", "5", "18")))
```

```{r variation_hedge_preparateData}
variationHedgeInd <- vali %>% 
  filter(studyversion == "hedge") %>% 
  mutate(
    ageInYearsExact = ifelse(sample == "kids", ageInMonths/12, 18), 
    ageInMonthsCentered = scale(ageInYearsExact, center = TRUE, scale = TRUE)[,1],
  ) %>% 
  group_by(sample, datacollection, ageInYears, ageInYearsExact, ageInMonthsCentered, subjID) %>%
  summarise(mean = mean(clickDistInBalloons, na.rm = TRUE))

@@

```{r age_hedge_model}
mDataAgeHedge <- vali %>% 
  filter(studyversion == "hedge" & sample == "kids") %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age = scale(ageInMonths, center = TRUE, scale = TRUE)[,1], 
  ) %>% 
  select(subjID, age, ageInMonths, datacollection, symmetricPosition, targetPosition, clickDistInBalloons)

mAgeHedge <- brm(clickDistInBalloons ~ age + datacollection + (symmetricPosition | subjID) + (1 | targetPosition),
                    data = mDataAgeHedge,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 100)

@@

```{r GLMM_plot_hedge}
nd_hedge <- mDataAgeHedge %>% 
  select(-c(clickDistInBalloons, symmetricPosition, targetPosition))

# extract model predictions for mock data
pred_hedge <- fitted(mAgeHedge, newdata = nd_hedge, re_formula = NA) %>% as_tibble() %>% bind_cols(mDataAgeHedge)

@@

  labs(x = "Age in years", y = "Average imprecision in target width", col = "Age group", shape = "Data collection mode", alpha = "Data collection mode", lty = "Data collection mode") + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 1)) + 
  scale_x_continuous(breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) +

@@

```{r variation_box_prepareData}
variationBoxInd <- vali %>% 
  filter(studyversion == "box") %>% 
  mutate(
    ageInYearsExact = ifelse(sample == "kids", ageInMonths/12, 18), 
    ageInMonthsCentered = scale(ageInYearsExact, center = TRUE, scale = TRUE)[,1],
  ) %>% 
  group_by(sample, datacollection, ageInYears, ageInYearsExact, ageInMonthsCentered, subjID) %>% 
  summarise(mean = mean(correctBox, na.rm = TRUE))

@@

```{r age_box_model}
# define data
mDataAgeBox <- vali %>% 
  filter(studyversion == "box" & sample == "kids") %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age = scale(ageInMonths, center = TRUE, scale = TRUE)[,1], 
  ) %>% 
  select(subjID, age, ageInMonths, datacollection, symmetricPosition, targetPosition, correctBox)

mAgeBox <- brm(correctBox ~ age + datacollection + (symmetricPosition | subjID) + (1 | targetPosition),
                    data = mDataAgeBox,
                    family = bernoulli(link = "logit"),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 100)

@@

```{r GLMM_plot_box}
# mAgeBox <- readRDS("../saves/mAgeBox.rds")

nd_box <- mDataAgeBox %>% 
  select(-c(correctBox, symmetricPosition, targetPosition))

@@

  labs(x = "Age in years", y = "Proportion correct", col = "Age in years", shape = "Data collection mode", alpha = "Data collection mode", lty = "Data collection mode") + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) + 
  scale_x_continuous(breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) + 
  theme(legend.position="none")

@@

```{r intCon_hedge_kids}
hedge_intCon_kids <- hedge_intCon %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_even_trials, y = mean_odd_trials)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 5, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  labs(x = "Even trials", y = "Odd trials")

@@

```{r intCon_box_kids}
box_intCon_kids <- box_intCon %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_even_trials, y = mean_odd_trials)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 1, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(x = "Even trials", y = "Odd trials")

@@

## Splithalf Reliability Table & Plot
```{r splithalf_plot}
splithalf_plot <- splithalf_data %>%
  mutate(
    sample = fct_rev(sample),
  ) %>%
  unite("samplegroup", c(studyversion, sample), sep = " - ", remove = FALSE) %>%
  mutate(samplegroup = fct_rev(samplegroup)) %>%
  unite("methodStrat", c(method, stratification), sep = " - ", remove = FALSE) %>%
  mutate(
    methodStrat = factor(methodStrat, levels = c("first-second - none", "first-second - target position", "odd-even - none", "odd-even - target position", "permutated - none", "permutated - target position", "Monte Carlo - none", "Monte Carlo - target position")) %>% fct_rev(),
  ) %>%
  ggplot(aes(x = methodStrat, y = coefficient, color = sample, group = studyversion)) +
    geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), colour = "black", width = .1, position = position_dodge(0.1)) +
    geom_point(position = position_dodge(0.1), size = 3) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    scale_color_manual(values = c("#006c66", "darkgrey"), guide = "none") +
    labs(x = "Splitting & stratification method", y = "") +
    coord_flip() +
    facet_wrap(~ samplegroup, nrow = 1) +
    theme(legend.position = "none", panel.spacing = unit(1.5, "lines"))

@@

# ggsave("../figures/tango_splithalf.png", width = 8, height = 2.5, scale = 1.2)
```

@@

```{r retest_hedge_kids}
hedge_retest_kids <- hedge_retest %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_reliday1, y = mean_reliday2)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 5, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  labs(x = "Testday 1", y = "Testday 2")

@@

```{r retest_box_kids}
box_retest_kids <- box_retest %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_reliday1, y = mean_reliday2)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#00b1ea", "#006c66", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 1, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = "Testday 1", y = "Testday 2")

@@

            # hedge: space between vari and reli
            nrow = 1, widths = c(1, 0.06, 1),
            labels = c("A", "", "")
  ),

@@

            # box: space between vari and reli
            nrow = 1, widths = c(1, 0.06, 1),
            labels = c("F", "", "")
  ),

@@

tango_arrangedplot
saveRDS(tango_arrangedplot, file = "../saves/tango-arrangedplot.rds")
# ggsave("../figures/tango_arrangedplot.png", width = 6.5, height = 5, scale = 1.6, bg = "white")
```

>>>>

    # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
    # the smaller the value, the more central the position
    symmetricPosition = ifelse(studyversion == "box", 
                               # for box version
                               case_when(
                                  targetPosition == "box3" ~ 0, 
                                  targetPosition == "box2" | targetPosition == "box4" ~ 1,
                                  targetPosition == "box1" | targetPosition == "box5" ~ 2), 
                               # for hedge version
                               targetCentralityX
                       ),

@@

```{r vali_data}
inddiff <- testtrials %>% 
  filter(studytype == "vali" & (ageInYears < 6 | ageInYears == 18)) %>% 
  mutate(ageInYears = factor(ageInYears, levels = c("3", "4", "5", "18")))
```

```{r variation_hedge_preparateData}
variationHedgeInd <- inddiff %>% 
  filter(studyversion == "hedge") %>% 
  mutate(
    ageInYearsExact = ifelse(sample == "kids", ageInMonths/12, 18), 
    ageInMonthsCentered = scale(ageInYearsExact, center = TRUE, scale = TRUE)[,1],
  ) %>% 
  group_by(sample, datacollection, ageInYears, ageInYearsExact, ageInMonthsCentered, subjID) %>%
  summarise(mean = mean(clickDistInBalloons, na.rm = TRUE))

@@

```{r age_hedge_model}
mDataAgeHedge <- inddiff %>% 
  filter(studyversion == "hedge" & sample == "kids") %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age = scale(ageInMonths), 
    symmetricPosition = scale(symmetricPosition),
    trialNr = trialNr - min(trialNr), 
  ) %>% 
  select(subjID, age, ageInMonths, datacollection, symmetricPosition, trialNr, clickDistInBalloons)

mAgeHedge <- brm(clickDistInBalloons ~ age + datacollection + symmetricPosition + trialNr + (1 + symmetricPosition + trialNr | subjID),
                    data = mDataAgeHedge,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4)

@@

```{r GLMM_plot_hedge}
nd_hedge <- mDataAgeHedge %>% 
  select(-c(clickDistInBalloons, trialNr)) %>% 
  mutate(trialNr = 0, symmetricPosition = 0)

pred_hedge <- fitted(mAgeHedge, newdata = nd_hedge, re_formula = NA) %>% as_tibble() %>% bind_cols(nd_hedge)

@@

  labs(x = "Age in years", y = "Average imprecision in target width", col = "Age group", shape = "Data collection mode", alpha = "Data collection mode", lty = "Data collection mode") + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 1)) + 
  scale_x_continuous(breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) +

@@

```{r variation_box_prepareData}
variationBoxInd <- inddiff %>% 
  filter(studyversion == "box") %>% 
  mutate(
    ageInYearsExact = ifelse(sample == "kids", ageInMonths/12, 18), 
    ageInMonthsCentered = scale(ageInYearsExact, center = TRUE, scale = TRUE)[,1],
  ) %>% 
  group_by(sample, datacollection, ageInYears, ageInYearsExact, ageInMonthsCentered, subjID) %>% 
  summarise(mean = mean(correctBox, na.rm = TRUE))

@@

```{r age_box_model}
# define data
mDataAgeBox <- inddiff %>% 
  filter(studyversion == "box" & sample == "kids") %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    age = scale(ageInMonths),
    symmetricPosition = scale(symmetricPosition),
    trialNr = trialNr - min(trialNr),
  ) %>% 
  select(subjID, age, ageInMonths, datacollection, symmetricPosition, trialNr, correctBox)

mAgeBox <- brm(correctBox ~ age + datacollection + symmetricPosition + trialNr + (1 + symmetricPosition + trialNr | subjID),
                    data = mDataAgeBox,
                    family = bernoulli(link = "logit"),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4)

@@

```{r GLMM_plot_box}
# mAgeBox <- readRDS("../saves/mAgeBox.rds")
nd_box <- mDataAgeBox %>% 
  select(-c(correctBox, trialNr)) %>% 
  mutate(trialNr = 0, symmetricPosition = 0)

@@

  labs(x = "Age in years", y = "Proportion correct", col = "Age in years", shape = "Data collection mode", alpha = "Data collection mode", lty = "Data collection mode") + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "darkgrey"), breaks = c("3", "4", "5", "18"), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) + 
  scale_x_continuous(breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) + 
  theme(legend.position="none")

@@

```{r intCon_hedge_kids}
hedge_intCon_kids <- hedge_intCon %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_even_trials, y = mean_odd_trials)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 5, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  labs(x = "Even trials", y = "Odd trials")

@@

```{r intCon_box_kids}
box_intCon_kids <- box_intCon %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_even_trials, y = mean_odd_trials)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 1, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = seq(0, 1, 0.2)) +
  labs(x = "Even trials", y = "Odd trials")

@@

## Splithalf Reliability Table & Plot
```{r splithalf_plot}
splithalf_plot <- splithalf_data %>%
  mutate(
    sample = factor(sample, levels = c("kids", "adults"), labels = c("children", "adults")),
  ) %>%
  unite("samplegroup", c(studyversion, sample), sep = " - ", remove = FALSE) %>%
  mutate(samplegroup = fct_rev(samplegroup)) %>%
  unite("methodStrat", c(method, stratification), sep = " - ", remove = FALSE) %>%
  mutate(
    methodStrat = factor(methodStrat, levels = c("first-second - none", "first-second - target position", "odd-even - none", "odd-even - target position", "permutated - none", "permutated - target position", "Monte Carlo - none", "Monte Carlo - target position")) %>% fct_rev(),
  ) %>%
  ggplot(aes(x = methodStrat, y = coefficient, color = sample, group = studyversion)) +
    geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), colour = "black", width = .1, position = position_dodge(0.1)) +
    geom_point(position = position_dodge(0.1), size = 3) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    scale_color_manual(values = c("#006c66", "darkgrey"), guide = "none") +
    labs(x = "Splitting & stratification method", y = "") +
    coord_flip() +
    facet_wrap(~ samplegroup, nrow = 1) +
    theme(legend.position = "none", panel.spacing = unit(1.5, "lines"))

@@

ggsave("../figures/tango_splithalf.png", width = 8, height = 2.5, scale = 1.2)
```

@@

```{r retest_hedge_kids}
hedge_retest_kids <- hedge_retest %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_reliday1, y = mean_reliday2)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 5, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1)) +
  labs(x = "Testday 1", y = "Testday 2")

@@

```{r retest_box_kids}
box_retest_kids <- box_retest %>% 
  filter(sample == "kids") %>% 
  ggplot(., aes(x = mean_reliday1, y = mean_reliday2)) + 
  geom_abline(intercept = 0, slope = 1, col = "darkgrey", linetype = "longdash") +
  geom_point(aes(col = ageInYears), size = 0.7) + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), guide = "none") +
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0, label.y = 1, aes(label = paste(..r.label..)), size = 3.5, r.accuracy = 0.01, cor.coef.name = "r") + 
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = "Testday 1", y = "Testday 2")

@@

            # hedge: space between inddiff and reli
            nrow = 1, widths = c(1, 0.06, 1),
            labels = c("A", "", "")
  ),

@@

            # box: space between inddiff and reli
            nrow = 1, widths = c(1, 0.06, 1),
            labels = c("F", "", "")
  ),

@@

tango_arrangedplot
saveRDS(tango_arrangedplot, file = "../saves/tango-arrangedplot.rds")
ggsave("../figures/tango_arrangedplot.png", width = 6.5, height = 5, scale = 1.6, bg = "white")
```
